\bf Creating optimal conditions for reproducible data analysis in R with \texttt{fertile}

### 2019-12-01 fertile.tex
more abstract drafting (.tex file)

In data science, an analyst has created \emph{reproducible} work if
their code and data produces identical results when given to another
analyst.

This idea of reproducibility is becoming increasingly important to the
advancement of our scientific knowledge. One reason for this is that
reproducible research encourages transparency. Providing one's audience
with well organized, reproducible code and results allows readers to
understand the steps taken to generate findings from raw data and
determine for themselves whether or not they believe the results to be
reliable or trustworthy.

Additionally, reproducibility encourages collaboration and extended
research, allowing others to easily apply the methods used in one
project to their own work with minimal effort.

@@

• Note: What are these behavioral and technical challenges, and which
ones should we mention?

@@

• Why focus on R?

>>>>

In the field of data science, an analyst has created \emph{reproducible}
work if: 1) their code and data files have been made publicly available,
and 2) these files produce identical results when run by another
analyst.

In the age of modern computer technology, where knowledge can be
instantly shared worldwide, reproducibility is becoming incredibly
important to the advancement of scientific knowledge.

The first reason for this is that reproducible research is transparent.
When researchers provide the code and data used for their work in a well
organized and reproducible format, readers are able to understand the
steps taken to generate findings from raw data and determine whether or
not they believe the results to be reliable or trustworthy. Without
reproducibility, research results must be blindly trusted.

The second reason is that reproducibility allows for collaboration and
extended research. The creators of reproducible research can easily
receive feedback and recommendations on their work, and others
interested in the research topic can see the code and easily apply the
methods and ideas used in one project to their own future work with
minimal effort.

@@

@@

Much of the discussion around reproducibility is also generalized,
written in a way applicable to users working with a variety of
statistical software programs. As a result, analysts interested in
practicing reproducibility cannot easily access recommendations tailored
specifically to their programming language or software of choice.

Additionally, all statistical software programs operate differently.
Since analysts do not all share the same software preferences, there can
be no single set of recommendations or guidelines for reproducibility.
Instead, recommendations must be specific to certain software programs.

In this paper, we focus on reproducibility in the R programming
language, with a specific concentration on the RStudio integrated
development environment, due to its popularity for statistical analyses
and the ease with which analysts can download and begin using the
software.

We propose an R package called \texttt{fertile}, which focuses on a
series of common reproducibility-harming mistakes programmers make while
conducting data science projects in R, warning users of their errors and
providing recommendations for how to correct them.

\texttt{fertile} operates in two modes: proactively (to prevent
reproducibility mistakes from happening in the first place), and
retroactively (analyzing code that is already written for potential
problems).

Retroactively, \texttt{fertile} is designed to be run on an R Project
folder, analyzing the project structure and the analyst's use of file
paths and considering randomness, among other areas of interest.
\texttt{fertile} creates reproducibility reports, identifying mistakes
that users have made and providing recommendations for remedies. For
example, fertile might identify an analyst's use of randomness in code
and recommend setting a seed.

Proactively, the package works similarly, identifying mistakes as they
are made by the user and outputting an educational warning message
identifying, and providing a solution for, the mistake. In its proactive
warning system, \texttt{fertile} focuses primarily on the use of file
paths. As users execute code, \texttt{fertile} looks for file paths
passed to functions, identifying when absolute paths are provided or
when paths point to a location outside of the project directory.

\texttt{fertile} is designed to educate the user about why the mistakes
are problematic, and how to fix them. It is written for a wide audience,
simple enough to be used by students in an introductory data science
course.
